Temporal and Geospatial Patterns Through Local Weather Simulation in BigPetStore

### 2015-01-01 bdse2015_paper/paper.tex
Update descriptions of model

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Adding Geospatial Patterns to the BigPetStore Data Generator with a Local Weather Model}

@@

The models were derived for daily average temperatures ($T(t)$, $^\circ$F), daily total precipitation ($P(t)$, in), and daily average wind speed ($V(t)$, mph). The temperature for each location is modeled as a first-order Fourier series plus noise generated by an Ornstein-Uhlenbeck process $Z(t)$ \cite{Gardiner09}:

@@

The daily average wind speed for each region is modeled using a first-order Fourier series and stochastic equation:

@@

The average temperature $a_0$, $a_1$, $a_2$, inverse daily average precipitation $\lambda$, average daily wind speed $b_0$, $b_1$, $b_2$, $k$, and $\theta$ are \textcolor{red}{fit to the observed data (how?)} from NOAA QCLCD for each zip code.  The damping coefficient $\gamma$ is determined empirically.

From the temperature, precipitation and wind speed, the snowfall ($S(t)$, in), rainfall ($R(T)$, in), and wind chill ($C(T)$, $^\circ$F) \textcolor{red}{cite} are calculated as follows:

@@

\begin{align}
C(T) = 35.74 + &0.6215T(t) - 35.75V^{0.16}(t) + \\ \nonumber
&0.4275T(t)V^{0.16}(t)
\end{align}

The snowfall, rainfall, and wind chill are used as input to a conditional probability density function (PDF) giving the probability of a transaction occuring on the date of time $t$ given the weather conditions:

\begin{equation}
P(\text{trans}|t, S(T), R(T), C(T)) = \dots
\end{equation}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{Adding Geospatial Patterns to the BigPetStore Data Generator Through Local Weather Simulation}
\title{Temporal and Geospatial Patterns Through Local Weather Simulation in BigPetStore}

@@

Models were derived for daily average temperatures ($T(t)$, $^\circ$F), daily total precipitation ($P(t)$, in), and daily average wind speed ($V(t)$, mph). The temperature for each location is modeled as a first-order Fourier series plus noise generated by an Ornstein-Uhlenbeck process $Z(t)$ \cite{Gardiner09}:

@@

where $\lambda$ is inverse of the daily average precipitation.

The daily average wind speed for each region is modeled using a first-order Fourier series and values sampled independently day-to-day from an Erlang distribution:

@@

where $k$ and $\theta$ are the shape and scale parameters, respectively.

From the temperature, precipitation and wind speed, the snowfall ($S(t)$, in), rainfall ($R(T)$, in), and wind chill ($WC(T)$, $^\circ$F) \textcolor{red}{cite} are calculated as follows:

@@

\begin{align}
WC(T) = 35.74 + &0.6215T(t) - 35.75V^{0.16}(t) \, + \\ \nonumber
&0.4275T(t)V^{0.16}(t)
\end{align}

Model parameters were estimated separately for each weather station in the NOAA QCLCD data from October 2011 to September 2014.  Weather stations missing more than 5\% of the data were excluded.  The Fourier coefficients $a_0$, $a_1$, $a_2$, $b_0$, $b_1$, and $b_2$ were found from Fourier transforms of the daily average temperature and daily average wind speed data. The distribution parameters $\sigma$, $\lambda$, $k$, and $\theta$ were computed directly from the data using the Maximum Likelihood Estimaters (MLE).  The damping coefficient $\gamma$ was determined empirically and the same value was used for every weather station.

Local weather simulations are run for each store using model parameters from the closest weather stations. The closest weather station is found for each store by computing the distances from the GPS coordinates of the stores' zip codes and weather station from the NOAA QCLCD data.

\subsection{Incorporating the Effect of Weather on Transaction Times}
The transaction time model was updated to incorporate the effect of the daily weather on customers' shopping times.

Separate conditional probability density functions (PDFs) giving the probability of a transaction occuring on day $d_i$ is given for the daily wind chill and snow fall.  The PDFs are modeled using sigmoidal functions:

\begin{equation}
P(\text{trans}_{d_i}=\text{YES}|x_{d_i}) = \frac{A}{1+\exp(-B(x-C))}+D
\end{equation}

\textcolor{red}{TODO table of parameters}

The probability of a transaction occurring on a day $d_i$ is given by the minimum probability of the individual PDFs:

\begin{align}
P(\text{trans}_{d_i}=\text{YES} | S_{d_i}, WC_{d_i}) = &\min \, \{ P(\text{trans}_{d_i}=\text{YES}|S_{d_i}), \nonumber \\ 
&P(\text{trans}_{d_i}=\text{YES} | WC_{d_i}) \, \}
\end{align}

The transaction time model was modified by adding the weather PDF to the transaction time acceptance probability PDF for transaction time $t_i$:

\begin{align}
P(\text{trans}_{t_i}&=\text{YES}|t_{i-1}, S_{d_i}, WC_{d_i}) = \nonumber \\
&Z^{-1} P(\text{trans}_{t_i}=\text{YES}|t_{i-1}) P(\text{trans}_{d_i}=\text{YES} | S_{d_i}, WC_{d_i})
\end{align}

where $Z$ is the normalization factor.
